---
title: "RDA fungi"
format:
  html:
    theme: spacelab
    toc: true
    number-sections: true
    digits: 3
---

```{r}
pacman::p_load(pacman, here, BiodiversityR, tictoc, tidyverse, V8)
```

# Load data for RDA fungi      

## Metadata
### Load Metadata - plant morphology - phenanthrene - qPCR
```{r Metadata - data sets loading}
meta.morpho.phe <- read.csv(file = here("data","meta.otu.june2020.txt"), dec = ".", header = T, sep = "\t", comment.char = "", row.names = 1) #load
meta.morpho.phe$code <- stringr::str_replace_all(meta.morpho.phe$code, '[_]', '.')
meta.morpho.phe.b.keep <- meta.morpho.phe$compartment == "bulk"
meta.morpho.phe.r.keep <- meta.morpho.phe$compartment == "rhizo"
meta.morpho.phe.b <- meta.morpho.phe[meta.morpho.phe.b.keep,]
meta.morpho.phe.r <- meta.morpho.phe[meta.morpho.phe.r.keep,]
meta.morpho.phe <- rbind(meta.morpho.phe.b,meta.morpho.phe.r) # keep only bulk and rhizosphere samples
```

### Recode variables
```{r Metadata - recode}
meta.morpho.phe$contamination <- as.factor(meta.morpho.phe$contamination)
meta.morpho.phe$compartment <- as.factor(meta.morpho.phe$compartment)
meta.morpho.phe$diversity_status <- as.factor(meta.morpho.phe$diversity_status)
meta.morpho.phe$diversity_status = factor(meta.morpho.phe$diversity_status, 
                                      levels = c("O", "C", "N", "E", "CN", "CE", "EN", "CEN"),
                                      labels = c("BF", "C", "N", "E", "CN", "CE", "EN", "CEN"))
```

### Keep only needed variables for the rda
```{r Metadata - keep variables}
meta.morpho.phe$ver <- NULL
meta.morpho.phe$col <- NULL
meta.morpho.phe$nem <- NULL
meta.morpho.phe$treatment <- NULL
meta.morpho.phe$cont.comp <- NULL
meta.morpho.phe$block <- NULL
meta.morpho.phe$pot.number <- NULL
meta.morpho.phe$pah.rhd.gp.c <- NULL
meta.morpho.phe$pah.rhd.gn.c <- NULL
meta.morpho.phe$branch <- NULL
meta.morpho.phe$fw_aerial = NULL
meta.morpho.phe$dw_aerial = NULL
meta.morpho.phe$dw_root = NULL
meta.morpho.phe$RSR = NULL
meta.morpho.phe$height = NULL
meta.morpho.phe$phe.b = NULL
meta.morpho.phe$phe.r = NULL
meta.morpho.phe$phe.total = NULL
meta.morpho.phe$pah.rhd.gp.b = NULL
meta.morpho.phe$pah.rhd.gp.r = NULL
meta.morpho.phe$pah.rhd.gn.b = NULL
meta.morpho.phe$pah.rhd.gn.r = NULL
```


### Tidy up
```{r Metadata - tidy up}
rm(meta.morpho.phe.b)
rm(meta.morpho.phe.r)
rm(meta.morpho.phe.b.keep)
rm(meta.morpho.phe.r.keep)
```

### Filter out incomplete cases
```{r}
meta.morpho.phe <- meta.morpho.phe[complete.cases(meta.morpho.phe), ] # Down from 192 to 184 cases #188?
```

### Test for autocorrelation of variables
```{r}
autocorrelations =  (cor(meta.morpho.phe[, c( 'biomass', 'phe', 'gn.copies.g.soil', 'gp.copies.g.soil')], method = "spearman"))
```

## Fungi

```{r Fungi - data sets loading}
#### ----------------- ITS - General community --------------------------------- 
obj1.rawits <- read.csv(file = here("data", "ITS_feature_table_filtered.tsv"), dec = ".", sep = "\t", header = T, row.names = 1, comment.char = "")
```

### Recode community
```{r Fungi - recode community}
samplenames <- colnames(obj1.rawits) #  sample names to be changed
goodsamplenames <- samplenames %-% '[X]'# erase the X added by R 
goodsamplenames <- stringr::str_replace_all(goodsamplenames, '[.]', '_') # code as in meta.morpho.phe
goodsamplenames # check did it work?
colnames(obj1.rawits) <- goodsamplenames #change old colnames for good colnames
dim(obj1.rawits) #check dimensions: 5408 obs of 193 variables
```

### Tidy up
```{r Fungi - Tidy intermediate objs}
rm(goodsamplenames)
rm(samplenames) # erases junk intermediate data to prevent future problems
```

# Curing community
## Taxonomy table
### Select taxonomy column
```{r Fungi - Curing taxonomy data tables}
taxonomy.its <- as.data.frame(obj1.rawits[,ncol(obj1.rawits)]) # taxonomy from last column as independent table
rownames(taxonomy.its) <- rownames((obj1.rawits)) # keep ASV numbers
check.datasets(taxonomy.its, obj1.rawits) # Check if numbers the same
rownames(obj1.rawits) <- paste("ASV",rownames(obj1.rawits),sep = "_")# modify ASV names
rownames(taxonomy.its) <- paste("ASV",rownames(taxonomy.its),sep = "_") # modify ASV names
check.datasets(taxonomy.its, obj1.rawits) # Check if numbers the same
```

### Separate taxonomic levels in taxonomy table
```{r Fungi - Curing taxonomy data tables}
colnames(taxonomy.its) <- c("Taxonomy") # Rename taxonomy column
# separate taxonomic groups in variables ----
tax.clean <- separate(taxonomy.its, Taxonomy, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species", "Strain"), sep = ";")
tax.clean <- tax.clean[,-which(names(tax.clean) %in% c("Species", "Strain"))]
tax.clean <- as.data.frame(tax.clean)
as.factor(tax.clean$Domain) # 1 Domain - see if there is contamination with chloroplast
as.factor(tax.clean$Phylum) # 14 Phyla 
as.factor(tax.clean$Class) # 41 classes
as.factor(tax.clean$Order) # 106 orders
as.factor(tax.clean$Family) # 261 Families
as.factor(tax.clean$Genus) # 599 genus
str(tax.clean)
```

### Eliminate uninformative characters
```{r Fungi - Curing taxonomy data tables}
# eliminate the characters one by one or all using stringr package -----
tax.clean$Domain <- stringr::str_replace(tax.clean$Domain,'[k]', '')
tax.clean$Domain <- stringr::str_replace_all(tax.clean$Domain,'[__]', '')
tax.clean$Phylum <- stringr::str_replace(tax.clean$Phylum,'[p__]', '')
tax.clean$Phylum <- stringr::str_replace_all(tax.clean$Phylum,'[__]', '')
tax.clean$Class <- stringr::str_replace(tax.clean$Class,'[c__]', '')
tax.clean$Class <- stringr::str_replace_all(tax.clean$Class,'[__]', '')
tax.clean$Order <- stringr::str_replace(tax.clean$Order, '[o]', '')
tax.clean$Order <- stringr::str_replace_all(tax.clean$Order, '[__]', '')
tax.clean$Family <- stringr::str_replace(tax.clean$Family, '[f]', '')
tax.clean$Family <- stringr::str_replace_all(tax.clean$Family, '[__]', '')
tax.clean$Genus <- stringr::str_replace(tax.clean$Genus, '[g]','')
tax.clean$Genus <- stringr::str_replace_all(tax.clean$Genus, '[__]', '')
str(tax.clean)
```

### Collate clean tax to community
```{r Fungi - Curing taxonomy data tables}
# collate otus with clean taxonomy table
obj1.rawits_bis <- cbind(obj1.rawits[,1:192], tax.clean)
# tyding up
rm(otuids.its)
rm(taxonomy.its)
```

## Community  matrix
### Transpose community matrix
```{r Fungi -  Preparing  abundance matrix, ordering tables}
tax.clean.its <- tax.clean # working file for taxonomy
comits <- obj1.rawits[,!names(obj1.rawits) %in% c('taxonomy')] # create community table with asv table with only abundances not taxonomy
comits.t <- t(comits) # transposing table, samples as rows, species as columns to match metadata
```
### Inner-join with meta.morpho.phe
Keep community samples that match environmental samples
```{r}
comits.t.filtered <- same.sites(comits.t, meta.morpho.phe)
check.datasets(comits.t.filtered, meta.morpho.phe) # check row names are the same
```

### Transform Hellinger
```{r}
comits.hell = disttransform(comits.t.filtered, method = 'hellinger')
```

# RDA steps
## Standardize 
```{r}
meta.z = meta.morpho.phe |> mutate_if(is.numeric,scale)
```

## Run RDA model
```{r}
#db-RDA Function dbrda does not provide species weights because it uses a distance matrix
comits.bray = vegdist(comits.hell, method = 'bray')
dbrda.its = dbrda(comits.bray ~ contamination + compartment + diversity_status + biomass + phe + gn.copies.g.soil + gp.copies.g.soil, data = meta.z, 
               scaling = "species")
RsquareAdj(dbrda.its)
anova.rda.its = anova.cca(dbrda.its, step = 999, by = "term")
summary(dbrda.its)
capture.output(anova.rda.its, file = here::here("output", "tables", "its", "anova-dbrda-its.txt"))
#tb-RDA use on community data after transformation: for example, hellinger transformed abundances
tbrda.its = rda(comits.hell ~ contamination + compartment + diversity_status + biomass + phe + gn.copies.g.soil + gp.copies.g.soil, data = meta.z, 
               scaling = "species")
#summary(rda.its1)
attach(meta.z)
anova.cca((tbrda.its))

#partial-RDA calculates rda with explanatory variables after considering the effect of covariates
# Alternative syntax for the partial RDA:
partial.rda.its <- rda(comits.hell ~ biomass + phe + gn.copies.g.soil + gp.copies.g.soil  + # these are the effects we are interested in
                          Condition(contamination + compartment + diversity_status), # these are the covariates
                       data = meta.z)
anova.cca(partial.rda.its)
summary(partial.rda.its)

```






# Ordiplot
```{r}
par(mfrow = c(2,2))
plot.its.partial = ordiplot(partial.rda.its, scaling = 2, choices = c(1,2))
plot.its.db = ordiplot(dbrda.its, scaling = 2, choices = c(1,2))
plot.its.tb = ordiplot(tbrda.its, scaling = 2, choices = c(1,2))
```

## Extract scores for ggplotting
```{r}
sites.long3 <- sites.long(plot.its.db, env.data = meta.morpho.phe)
tic()
spec.envfit <- envfit(plot.its.db, env = comits.hell, permutations = 1) #Repeat this when at home for 99 permutations for around 20-30 minutes
toc()
spec.data.envfit <- data.frame(r = spec.envfit$vectors$r, p = spec.envfit$vectors$pvals)
species.long2 <- species.long(plot.its.db, spec.data = spec.data.envfit)
species.long3 <- species.long2[species.long2$r >= 0.35, ]
species.long3$labels <- make.cepnames(species.long3$labels)
dim(species.long3)
# Which ASVs were retained?
its.tax.retained =  merge(tax.clean, species.long3, by = 0)

vectors.envfit <- envfit(plot.its.db, env = meta.morpho.phe)
vectors.long3 <- vectorfit.long(vectors.envfit)
vectors.long3

axis.long3 <- axis.long(dbrda.its, choices=c(1, 2))
axis.long3

phe.surface <- ordisurf(plot.its.db, y = phe)
phe.surface
phe.grid <- ordisurfgrid.long(phe.surface)

#biomass <- ordisurf(plot.its.db, y = biomass)
#biomass
#biomass.grid <- ordisurfgrid.long(biomass)
```


## Extract scores for ggplotting
```{r}
sites.long3 <- sites.long(plot.its, env.data = meta.morpho.phe)
tic()
spec.envfit <- envfit(plot.its, env= comits.hell, permutations=1) #Repeat this when at home for 99 permutations for around 20-30 minutes
toc()
spec.data.envfit <- data.frame(r = spec.envfit$vectors$r, p = spec.envfit$vectors$pvals)
species.long2 <- species.long(plot.its, spec.data = spec.data.envfit)
species.long3 <- species.long2[species.long2$r >= 0.35, ]
species.long3$labels <- make.cepnames(species.long3$labels)
dim(species.long3)
# Which ASVs were retained?
its.tax.retained =  merge(tax.clean, species.long3, by=0)

vectors.envfit <- envfit(plot.its, env=meta.morpho.phe)
vectors.long3 <- vectorfit.long(vectors.envfit)
vectors.long3

axis.long3 <- axis.long(rda.its1, choices=c(1, 2))
axis.long3

phe.surface <- ordisurf(plot.its, y = phe)
phe.surface
phe.grid <- ordisurfgrid.long(phe.surface)

#biomass <- ordisurf(plot.its, y = biomass)
#biomass
#biomass.grid <- ordisurfgrid.long(biomass)
```


# RDA PLOT
## Create ggplot theme
```{r}
BioR.theme <- theme(
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid = element_blank(),
        axis.line = element_line("gray25", linewidth = 1),
        text = element_text(size = 10),
        axis.text = element_text(size = 10, colour = "gray25"),
        axis.title = element_text(size = 14, colour = "gray25", face = "bold"),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 14),
        legend.key = element_blank(),
        legend.position = "bottom")

```


```{r}
sites.long3$diversity_status = factor(sites.long3$diversity_status, 
                                      levels = c("O", "C", "N", "E", "CN", "CE", "EN", "CEN"),
                                      labels = c("BF", "C", "N", "E", "CN", "CE", "EN", "CEN"))
levels(sites.long3$diversity_status)
plotgg5 <- ggplot() + 
    geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
    geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
    xlab(axis.long3[1, "label"]) +
    ylab(axis.long3[2, "label"]) +  
    scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
    scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
    #geom_mark_hull(data=sites.long3,  aes(x=axis1, y=axis2, colour=diversity_status,   fill=after_scale(alpha(colour, 0.2))),   concavity=0.1, size=0.2, show.legend=FALSE) +
    geom_point(data=sites.long3, 
               aes(x=axis1, y=axis2, color=diversity_status, shape=compartment), 
               alpha=0.7, size=3) +
           scale_color_manual(" ", 
                           values = c( "BF" = "gray77", 
                                       "C" = "#edf8b1", 
                                       "E" = "#7fcdbb", 
                                       "N" = "#c7e9b4",
                                       "CE" = "#1d91c0", 
                                       "CN" = "#41b6c4", 
                                       "EN" = "#225ea8", 
                                       "CEN" = "#0c2c84"
                                      ) 
                           ) +
    geom_segment(data=species.long3, 
                 aes(x=0, y=0, xend=axis1*6, yend=axis2*6), 
                 colour="red", size=0.7, arrow=arrow()) +
    geom_text_repel(data=species.long3, 
                    aes(x=axis1*6, y=axis2*6, label=labels),
                    colour="black") +
    geom_segment(data=subset(vectors.long3, vector %in% vectors),
                 aes(x=0, y=0, xend=axis1*1, yend=axis2*1), 
                 colour="blue", size=0.7, arrow=arrow()) +
    geom_text_repel(data=subset(vectors.long3, vector %in% vectors), 
                    aes(x=axis1*1, y=axis2*1, label=vector),
                    colour="black") +
    BioR.theme +
    #ggsci::scale_colour_npg() +
    labs(shape="compartment", colour="diversity_status", fill="diversity_status") +
    coord_fixed(ratio=1)
plotgg5

# Keep only: gp, gn, phe and biomass!!
```





















